---
title: "Assignment_2"
author: "Vijay Charan Reddy Gottam"
date: "2023-09-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary


1. This new customer would be classified as 0, does not take the personal loan.
2. The best K is 3 that balances between overfitting and ignoring the predictor
   information.
3. Using the best K, the confusion matrix for the validation data is provided below with values of TP= 142, TN= 1786, FP= 63, FN=    9 with accuracy of 0.964 and other parameters.
4. After using the best K, the customer would be classified as 0, does not take the personal loan.


## Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer demographic information (age, income, etc.), the customer’s relationship with the bank (mortgage, securities account, etc.), and the customer response to the last personal loan campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the personal loan that was offered to them in the earlier campaign.

Partition the data into training (60%) and validation (40%) sets.

***

### Data Import and Cleaning

First, load the required libraries

```{r}
library(class)
library(caret)
library(e1071)
```
Read the data.

```{r }
universal.df <- read.csv("UniversalBank.csv")
dim(universal.df)
t(t(names(universal.df))) # The t function creates a transpose of the dataframe
```
Drop ID and ZIP
```{r}
universal.df <- universal.df[,-c(1,5)]
```

Split Data into 60% training and 40% validation. There are many ways to do this. We will look at 2 different ways. Before we split, let us transform categorical variables into dummy variables

```{r}
# Only Education needs to be converted to factor
universal.df$Education <- as.factor(universal.df$Education)

# Now, convert Education to Dummy Variables

groups <- dummyVars(~., data = universal.df) # This creates the dummy groups
universal_m.df <- as.data.frame(predict(groups,universal.df))


set.seed(1)  # Important to ensure that we get the same sample if we rerun the code
train.index <- sample(row.names(universal_m.df), 0.6*dim(universal_m.df)[1])
valid.index <- setdiff(row.names(universal_m.df), train.index)  
train.df <- universal_m.df[train.index,]
valid.df <- universal_m.df[valid.index,]
t(t(names(train.df)))

#Second approach

library(caTools)
set.seed(1)
split <- sample.split(universal_m.df, SplitRatio = 0.6)
training_set <- subset(universal_m.df, split == TRUE)
validation_set <- subset(universal_m.df, split == FALSE)

# Print the sizes of the training and validation sets
print(paste("The size of the training set is:", nrow(training_set)))
print(paste("The size of the validation set is:", nrow(validation_set)))
```

Now, let us normalize the data
```{r}
train.norm.df <- train.df[,-10] # Note that Personal Income is the 10th variable
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
```

### Questions

Consider the following customer:

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

```{r}
# We have converted all categorical variables to dummy variables
# Let's create a new sample
new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer
new.cust.norm <- new_customer
new.cust.norm <- predict(norm.values, new.cust.norm)

```

Now, let us predict using knn
```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1

```

***

2. What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider

accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, 
                                       as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.df[,2] == max(accuracy.df[,2])) 

```
```{r}
plot(accuracy.df$k,accuracy.df$overallaccuracy)
```



3. Show the confusion matrix for the validation data that results from using the best k.


```{r}
best_k <- which(accuracy.df$overallaccuracy == max(accuracy.df$overallaccuracy))

# Using the best k to predict on validation data
best_k_pred <- class::knn(train = train.norm.df, 
                          test = valid.norm.df, 
                          cl = train.df$Personal.Loan, 
                          k = best_k)

# Creating the confusion matrix
confusion_matrix <- confusionMatrix(best_k_pred, 
                                    as.factor(valid.df$Personal.Loan), 
                                    positive = "1")

# Displaying the confusion matrix
confusion_matrix

```
***


4. Consider the following customer: Age = 40, Experience = 10, Income = 84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 =1, Education_3 = 0,Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.

```{r}
# Assuming we have already calculated the best_k
new_customer_1 <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer
new.cust.norm <- new_customer_1
new.cust.norm<- predict(norm.values, new.cust.norm)

# Use the best_k to predict the classification
best_k_classification <- class::knn(train = train.norm.df, 
                                     test = new.cust.norm, 
                                     cl = train.df$Personal.Loan, 
                                     k = best_k)

# Display the classification result
best_k_classification

```



